---
title: "Extraction des données paysagères autour des sites landus"
author: "Jeremy Froidevaux"
date: "2024-12-01"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/jeremyfroidevaux/Library/CloudStorage/GoogleDrive-jfroidevaux@creamontblanc.org/Drive partagés/prototool/herbiland/analyses/landus")
rm(list=ls()) 
```

## Load libraries
```{r}
library <- c("terra","ggplot2","grid","gridExtra","cowplot","landscapemetrics","data.table","maps",
             "ggsn","dplyr","rworldmap","reshape2","sf","terra","tidyverse","purrr","openxlsx")

lapply(library, require, character.only = TRUE) 
```

## Data manipulation with raster and shape files
### Open CameraTrap csv file (and landus csv file if needed)
```{r}
releves = read.xlsx("data/releve_vegetation/saisieRELEVES_gpesFct_20220914_final.xlsx", sheet = "SURVEYS")
head(releves)
#summary(factor(landus$ref_project))
```

### Transformation of latlong in vector and georeferenced data vizualization
```{r}
#All data
lonlat<-cbind(releves$Longitude,releves$Latitude)
crdref <- "+proj=longlat +datum=WGS84"
pts <- vect(lonlat, crs=crdref)
pts
pts$id<-releves$plot_project
class(pts)
outfile <- "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Landus_placette_veg/placetteveg_locWGS84.shp"
writeVector(pts, outfile, overwrite=TRUE)


dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Landus_placette_veg"  
name.points = "placetteveg_locWGS84"
pts<-st_read(dsn = dir.posPoints, layer = name.points)
loc <- st_transform(pts,crs=32632)
dir.posEnvironmentalLayers1 = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif"
orionrast1<-rast(dir.posEnvironmentalLayers1)
plot(orionrast1)
points(loc)
outfile <- "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Landus_placette_veg/placetteveg_locL93.shp"
st_write(loc, outfile, overwrite=TRUE)

```

## Data extraction
### Part 1: Extract landuse proportion from ORION at local and landscape scales

```{r}
library(lavaan)
root <- "/Users/jeremyfroidevaux/Documents/Herbiland/CameraTrap/Analysis/Functions/" 
source(paste(root,"ExtractBufferRaster_KB_JF.r",sep="")) 

extractLandUseRaster( dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Landus_placette_veg",  
                                name.points = "placetteveg_locL93",
                                bw = c(20,250), #ajouter 20 m - 250 m  
                                id = "id",  
                                dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif",  
                                WorkingDirectory = "/Users/jeremyfroidevaux/Documents/Herbiland/Landus/Sites_description/Data/") 

```

### Part 2: Extract data on landscape configurational and compositional diversity using ORION

```{r}
dir.posPoints = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/Landus_placette_veg"  
name.points = "placetteveg_locL93"
loc<-st_read(dsn = dir.posPoints, layer = name.points)
loc <- st_transform(loc,crs=32632)
dir.posEnvironmentalLayers = "/Users/jeremyfroidevaux/Documents/Herbiland/GIS/ORION/LC_pred_CCVCMB_2806023_FINAL.tif"
orionrast<-rast(dir.posEnvironmentalLayers) 
orionrast
loc
plot(orionrast)
points(loc)
list<-list_lsm() #Choose landscape metric 

plm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 20, level= "landscape",metric=c("shdi","ed")) 
plm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 250, level= "landscape",metric=c("shdi","ed")) 
pcm20 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 20, level= "class",metric=c("ed")) 
pcm250 <- sample_lsm(orionrast, y = loc, shape = "circle", plot_id = loc$id,  
                    size = 250, level= "class",metric=c("ed")) 

plm20$scale<-20 
plm250$scale<-250
pcm20$scale<-20 
pcm250$scale<-250

pcm20sel<-pcm20[pcm20$class %in% c("5"), ]
pcm250sel<-pcm250[pcm250$class %in% c("5"), ]
pcm20sel$metric[pcm20sel$metric == 'ed'] <- 'edl'
pcm250sel$metric[pcm250sel$metric == 'ed'] <- 'edl'

df_list <- list(plm20,plm250,pcm20sel,pcm250sel)
mplm<-Reduce(function(x, y) merge(x, y, all=TRUE), df_list)

div <- dcast(mplm, plot_id + scale~ metric, fun.aggregate = sum)

#ggplot(subset(rmplm, scale %in% 50))+
#  geom_histogram(aes(x=ed))

div20<-subset(div,scale=="20")
div20<-div20 %>% rename_with(~paste0(.x,"20"), c(ed,edl,shdi ))

div250<-subset(div,scale=="250")
div250<-div250 %>% rename_with(~paste0(.x,"250"), c(ed,edl,shdi ))

df_list <- list(div20, div250)
divf<-df_list %>% reduce(full_join, by='plot_id')
divf<-divf %>% dplyr::select(-contains("scale"))
colnames(divf)[colnames(divf) == "plot_id"] <- "id"

write.csv(divf, "/Users/jeremyfroidevaux/Documents/Herbiland/Landus/Sites_description/Data/ORION_landscapeMetric.csv", row.names=FALSE) 

```

## Merge dataset
### ORION
```{r}
releves = read.xlsx("data/releve_vegetation/saisieRELEVES_gpesFct_20220914_final.xlsx", sheet = "SURVEYS")
releves$id<-releves$plot_project
orion<-fread("output/Orion_landuse.csv")
div<-fread("output/ORION_landscapeMetric.csv")

colnames(orion)<-gsub("1_", "prairieM", colnames(orion))
colnames(orion)<-gsub("2_", "foret", colnames(orion))
colnames(orion)<-gsub("3_", "limForet", colnames(orion))
colnames(orion)<-gsub("4_", "prairieS", colnames(orion))
colnames(orion)<-gsub("5_", "lande", colnames(orion))
colnames(orion)<-gsub("6_", "ecotoneLP", colnames(orion))
colnames(orion)<-gsub("7_", "prairieA", colnames(orion))
colnames(orion)<-gsub("8_", "rochersolnu", colnames(orion))
colnames(orion)<-gsub("9_", "glacierneige", colnames(orion))
colnames(orion)<-gsub("10_", "eau", colnames(orion))
colnames(orion)<-gsub("0_", "urbain", colnames(orion))

df0<-merge(releves,div,by="id")
df<-merge(df0,orion,by="id")

df$landeEl20<-df$lande20+df$limForet20+df$ecotoneLP20
df$landeEl250<-df$lande250+df$limForet250+df$ecotoneLP250

write.csv(df,"output/landus_releves_Paysage_ORION.csv")
```
